---
editor_options: 
  chunk_output_type: inline
---

# prep session

## load libs
```{r}
#'load libs
source("libs.R")
source("ggplot.opts.r")
```

## load full data

```{r}
#'load as rdata
load("cvore.full.rdata")
utm43n = "+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs"

library(sf)
#'make spatial, get ndvi
cvore.sf = st_as_sf(cvore.rev, coords = c("xutm", "yutm"))
st_crs(cvore.sf) = utm43n

#'subeset with only id and make spatial
cvore.sp = as(cvore.sf %>% dplyr::select(id), "Spatial")

```

## load background

```{r}
source("libs.spatial.r")
#'load evi raster layer
cvore.ndvi = raster("cvore.ndvi.tif")
cvore.landuse = raster("cvore.landuse.tif")

#'get ndvi for cvores
cvore.rev$ndvi = extract(cvore.ndvi, cvore.sp)

env = data.frame(ndvi = as.vector(cvore.ndvi), landuse = as.vector(cvore.landuse))

#'get factors for landuse
landuse.class = bind_rows(levels(cvore.landuse))
#'left join with cvore.rev
cvore = left_join(cvore.rev, landuse.class, by = c("landuse"="ID"))
#'leftjoin with env
env = env %>% left_join(., landuse.class, by = c("landuse"="ID"))
env$env = "env"

#'make species common names
cvore$sp = mapvalues(cvore$sp, c("Felis chaus","Vulpes bengalensis","Canis aureus"), c("Jungle cat","Indian fox","Golden jackal"))
```

## colour palette

```{r}
vcols3 = viridis(10)[c(2,6,10)]
```


## fixes per month

```{r}
#png(filename = "fix.heatmap.day.#png", height = 1600, width = 2000, res = 400)

count(cvore, id,day = round(as.numeric(julian(cvore$time, origin = as.POSIXct("2017-01-01",tz="IST"))))) %>% 
  ggplot()+
  geom_tile(aes(x = id, y = day, fill = n))+
  g2+
  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 8, family = "monospace"), legend.position = "top", legend.text = element_text(size = 6))+
 scale_fill_gradientn(colours = rev(brewer.pal(9, "RdBu")))+
  labs(list(x = "Animal",y = "2017 days",fill = "Fixes"))

#dev.off()
```


## density plot of available and used evi

```{r}
#'plot densities

#cairo_pdf(filename = "fig3ndvi_avail_used.pdf", height = 5, width = 5)
#png(filename = "ndvi_avail_used.#png", height = 1600, width = 2000, res = 400)
#'plot
ggplot()+
  geom_density(data = cvore, aes(x = ndvi, fill = sp), position = "identity", alpha = 0.6, col = "transparent")+
  scale_fill_manual(values = vcols3)+
  stat_density(data = env, aes(x = ndvi), position = "identity", lwd = 0.5, geom = "line",col = "grey20")+
  labs(list(x = "NDVI", y = "Probability density", fill = "Spp."))+
  g2+
  theme(legend.position = "top", legend.text = element_text(face = "italic"))

#dev.off()
```

## landuse and ndvi

```{r}
#'landuse classes
ggplot()+
  geom_violin(data = env, aes(x = category, y = ndvi))+
  labs(list(x = "Landuse",y = "NDVI"))+
  g2
```

## used and available landuse

```{r}
#'used and available landuse classes
ggplot()+
  geom_bar(data = cvore, aes(x = sp, fill = category), position = "fill")+
  geom_bar(data = env, aes(x = env, fill = category), position = "fill")+
  scale_fill_manual(values = c(terrain.colors(5)[1:4],"royalblue"))+
  g2+
  theme(legend.position = "top", axis.text.x = element_text(face="italic"))+
  labs(list(x= NULL, y = "Proportion", fill = "Landuse"))
```

# relate revisits, residence, and speed to landuse

## residence distributions

```{r}
#'density plot residence
ggplot()+
  stat_density(data = cvore, aes(x = residence, col = sp), position = "identity", alpha = 1, geom = "line")+
  scale_colour_manual(values = vcols3)+
  #stat_density(data = env, aes(x = ndvi), position = "identity", lwd = 0.5, geom = "line",col = "grey20")+
  labs(list(x = "Residence time (hrs)", y = "Probability density", fill = "Spp."))+
  xlim(0,1000)+
  g2+
  theme(legend.position = "top", legend.text = element_text(face = "italic"))
```

## residence and landuse

```{r}
#'plot boxplots of revisits, residence, speed
#png(filename = "residence.landuse.#png", height = 1600, width = 2000, res = 400)

ggplot()+
  geom_boxplot(data = cvore, aes(x = category, y = residence, fill = sp),col = 1, position = position_dodge(width = 0.7), lwd = 0.3, outlier.size = 0.2)+
  scale_fill_manual(values = viridis(10)[c(2,6,10)])+
 # scale_colour_manual(values = viridis(3))+
  coord_cartesian(ylim = c(0,2e3))+
  labs(list(x = "Landuse",y = "Residence time (hrs)"))+
  g1

#dev.off()
```

## revisit distributions

```{r}
#'density plot revisits
ggplot()+
  stat_density(data = cvore, aes(x = revisits, fill = sp), position = "identity", alpha = .6)+
  scale_fill_manual(values = vcols3)+
  #stat_density(data = env, aes(x = ndvi), position = "identity", lwd = 0.5, geom = "line",col = "grey20")+
  labs(list(x = "Residence time (hrs)", y = "Probability density", fill = "Spp."))+
 # xlim(0,1000)+
  g2+
  theme(legend.position = "top", legend.text = element_text(face = "italic"))
```

## revisits and landuse

```{r}
#png(filename = "revisits.landuse.#png", height = 1600, width = 2000, res = 400)

ggplot()+
  geom_boxplot(data = cvore, aes(x = category, y = revisits, fill = sp),col = 1, position = position_dodge(width = 0.7), lwd = 0.3, outlier.size = 0.2)+
  scale_fill_manual(values = viridis(10)[c(2,6,10)])+
 # scale_colour_manual(values = viridis(3))+
  coord_cartesian(ylim = c(0,5e2))+
  labs(list(x = "Landuse",y = "# revisits"))+
  g1

#dev.off()
```

## speed distribution

```{r}
#'density plot speed
#png(filename = "distribution.speed.#png", height = 1600, width = 2000, res = 400)

ggplot()+
  stat_density(data = cvore, aes(x = speed*3600, col = sp), position = "identity", geom="line", lwd = 0.6)+
  scale_colour_manual(values = vcols3)+
  #stat_density(data = env, aes(x = ndvi), position = "identity", lwd = 0.5, geom = "line",col = "grey20")+
  labs(list(x = "Speed (m/hr)", y = "Probability density", colour = "Spp."))+
  xlim(0,4000)+
  g2+
  theme(legend.position = "top", legend.text = element_text(face = "italic"))

#dev.off()
```

## speed and landuse

```{r}
#png(filename = "speed.landuse.#png", height = 1600, width = 2000, res = 400)

ggplot()+
  geom_boxplot(data = cvore, aes(x = category, y = speed*3600, fill = sp),col = 1, position = position_dodge(width = 0.7), lwd = 0.3, outlier.size = 0.2)+
  scale_fill_manual(values = viridis(10)[c(3,6,10)])+
 # scale_colour_manual(values = viridis(3))+
  coord_cartesian(ylim = c(0,5000))+
  labs(list(x = "Landuse", y = "Speed (m/hr)"))+
  g1

#dev.off()
```

# move object

```{r}
library(move)
#'arrange by id, time
cvore = cvore %>% arrange(id, time) %>% filter(year == 2017)
utm43n = "+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs"

#'convert to a move object by id and get basic move parameters
cvore.move = move(x = cvore$xutm, y = cvore$yutm,time = cvore$time, proj = utm43n, animal = cvore$id, data = cvore)

#'split the movestack and get lag, speed, angle
cvore.move = split(cvore.move)
#'initiate loop
for(i in 1:length(cvore.move)){
  x = cvore.move[[i]]
  cvore.move[[i]]$lag = as.numeric(c(timeLag(x), units = "mins"))
  cvore.move[[i]]$distance = c(distance(x),NA)
}

#'make df
cvore = cvore.move %>% 
  lapply(., function(x) cbind(x@idData, x@data)) %>% 
  bind_rows() %>% 
  arrange(id, time)

```

# distribution of lags by species

```{r}
#'get lag quantiles
cvore %>% dlply("sp") %>% lapply(., function(x) quantile(x$lag, na.rm=T, probs = seq(0,1,0.05)))

#png(filename = "lag.density.#png", height = 1600, width = 2000, res = 400)

#'plot lag dist density plots, remove all lags above 360 (6 hours)
ggplot()+
  stat_density(data = cvore %>% filter(lag <= 360), aes(x = lag), geom = "line")+
  labs(list(x = "Fix interval (mins)"))+
  facet_wrap(~sp)+
  g2+
  theme(strip.background=element_blank(), strip.text = element_text(colour = 1, face = "italic"))

#dev.off()
```

## timelag vs hour of day

```{r}
#'plot lag dist density plots, remove all lags above 360 (6 hours)
#png(filename = "density.hour.#png", height = 1600, width = 2000, res = 400)

cvore %>% filter(lag <= 360) %>% group_by(sp, hour) %>% summarise(lagm = mean(lag, na.rm = T), sd = sd(lag, na.rm = T), n = length(lag)) %>% mutate(ci = qnorm(0.975)*sd/sqrt(n)) %>% 

ggplot()+
  geom_pointrange(aes(x = hour, y = lagm, ymin = lagm-ci, ymax=lagm+ci), lwd = 0.2, size = 0.2)+
  labs(list(x = "Hour of day", y = "Fix interval (mins)"))+
  facet_wrap(~sp)+
  g2+
  theme(strip.background=element_blank(), strip.text = element_text(colour = 1, face = "italic"))

#dev.off()
```

# effect of ndvi, distance to settlements, and overlap

## prep data

```{r}
#'residence, revisits, speed at ndvi: melt the data, with closely spaced means for ndvi etc
cvore.melt = cvore %>% filter(lag <=360) %>% group_by(ndvi = round_any(ndvi, 0.05), settledist = round_any(settledist, 100), overlap) %>%
  melt(id.vars = c("sp","id","ndvi","settledist","overlap"), measure.vars = c("speed","lag","revisits","residence"))

#'get mean values for ndvi
cvore.melt.ndvi = cvore.melt %>% group_by(sp,ndvi,variable) %>% summarise(m = mean(value), sd = sd(value), n = length(value)) %>% mutate(ci = qnorm(0.975)*sd/sqrt(n)) 

#'mean values for settlement distance
cvore.melt.settledist = cvore.melt %>% group_by(sp,settledist,variable) %>% summarise(m = mean(value), sd = sd(value), n = length(value)) %>% mutate(ci = qnorm(0.975)*sd/sqrt(n))

#'mean values for overlap
cvore.melt.overlap = cvore.melt %>% group_by(sp,overlap,variable) %>% summarise(m = mean(value), sd = sd(value), n = length(value)) %>% mutate(ci = qnorm(0.975)*sd/sqrt(n))
```

## plot data

```{r}
#'plot
#png(filename = "movevars.ndvi.#png", height = 1600, width = 2000, res = 400)
#'mean values ~ ndvi ~ sp
ggplot()+
  geom_pointrange(data = cvore.melt.ndvi, aes(x = ndvi, y = m, ymin = m-ci, ymax = m+ci, fill = sp), lwd = 0.2, size = 3, pch = 21, stroke = 0.2, fatten = 12)+
  scale_fill_manual(values = viridis(3))+
  facet_grid(variable~sp, scales = "free")+
  labs(list(x = "NDVI", y = "Value"))+
  g1+
  theme(strip.text = element_text(colour=1))

#dev.off()
```

## plot data

```{r}
#
#png(filename = "movevars.settlement.#png", height = 1600, width = 2000, res = 400)
#'mean values ~ settledist ~ sp
ggplot()+
  geom_pointrange(data = cvore.melt.settledist, aes(x = settledist, y = m, ymin = m-ci, ymax = m+ci, fill = sp), lwd = 0.2, size = 3, pch = 21, stroke = 0.2, fatten = 12)+
  scale_fill_manual(values = viridis(3))+
  facet_grid(variable~sp, scales = "free")+
  labs(list(x = "Distance to settlement (m)", y = "Value"))+
  xlim(0,2000)+
  g1+
  theme(strip.text = element_text(colour=1))

#dev.off()
```

## plot data

```{r}
#png(filename = "movevars.overlap.#png", height = 1600, width = 2000, res = 400)

#'mean values ~ overlap ~ sp
ggplot()+
  geom_violin(data = cvore.melt %>% filter(!is.na(overlap)), aes(x = as.factor(overlap), y = value), lwd = 0.2, size = 3, pch = 21, stroke = 0.2, fatten = 12)+
  scale_fill_manual(values = viridis(3))+
  facet_grid(variable~sp, scales = "free", drop = T)+
  labs(list(x = "Overlap #", y = "Value"))+
  g1+
  theme(strip.text = element_text(colour=1))

#dev.off()
```


